### 2019-11-20 paper/rebuttal.txt
First update to answer referees

Dear Dr. van der Spoel:

Thank you for submitting your manuscript to Journal of Chemical
 Information and Modeling. Please find the enclosed reviews of your
 manuscript. The reviewers have found significant problems with the
 manuscript in its current form, but we are willing to consider a
 revised manuscript in which you have made major revisions and
 responded to the reviewers' concerns. If you resubmit a revised
 manuscript, there may be a second round of external review, perhaps
 including new reviewers.

We would like to receive your revision as soon as possible, by
17-Jan-2020 at the latest.

In addition to addressing the reviewers' comments, please make each of the
technical corrections listed below:

MANUSCRIPT
- Please remove "et al." from your references and list all authors' names.
ANSWER:

- Update the publication information in Reference 56 and Reference 73 when
available.
ANSWER:

- Please include a Table of Contents graphic on the very last page of
your manuscript.  This graphic, in conjunction with the manuscript
title, should give the reader a summary of the research described in
the paper. The Table of Contents graphic will also be published in the
Abstract of the manuscript. This graphic can be no wider than 8 cm and
no taller than 4 cm and should be prepared following the
specifications under Artwork. Text should be limited to essential
technical labels.  More guidance is available at
http://pubs.acs.org/paragonplus/submission/toc_abstract_graphics_guidelines.pdf.
ANSWER:

- The title page of your manuscript file needs more complete
 affiliation information for all authors. Could you please provide
 updated affiliations including all of the following information
 (where possible)? Department and institute (or company name), city,
 state/region, postal code, country.  Please indicate the
 corresponding author's email address with an asterisk.
ANSWER:

- Funding Sources: Authors are required to report ALL funding sources
  and grant/award numbers relevant to this manuscript. Enter all
  sources of funding for ALL authors relevant to this manuscript in
  BOTH the Open Funder Registry tool in ACS Paragon Plus and in the
  manuscript to meet this requirement. See
  http://pubs.acs.org/page/4authors/funder_options.html for complete
  instructions.
ANSWER:

- ORCID: Authors submitting manuscript revisions are required to
  provide their own validated ORCID iDs before completing the
  submission, if an ORCID iD is not already associated with their ACS
  Paragon Plus user profiles. This iD may be provided during original
  manuscript submission or when submitting the manuscript
  revision. You can provide only your own ORCID iD, a unique
  researcher identifier. If your ORCID iD is not already validated and
  associated with your ACS Paragon Plus user profile, you may do so by
  following the ORCID-related links in the Email/Name section of your
  ACS Paragon Plus account. All authors are encouraged to register for
  and associate their own ORCID iDs with their ACS Paragon Plus
  profiles. The ORCID iD will be displayed in the published article
  for any author on a manuscript who has a validated ORCID iD
  associated with ACS Paragon Plus when the manuscript is
  accepted. Learn more at http://www.orcid.org.
ANSWER:

- On resubmission you must submit the following:
a) A point-by-point response to the reviewers.
b) A clean, unmarked copy of the manuscript document file.
c) A marked copy of the revised manuscript in which all the changes are clearly tracked or highlighted. Upload this file as Supporting Information for Review.
d) A clean, unmarked Supporting Information for Publication file, without any highlighting or tracked changes.

- The reviewers indicate that the quality of the written English could
  be improved. Please consider using a native English speaker or an
  English-language editing service to assist with this. The ACS
  maintains a list of editing services:
  https://authoringservices.acs.org/.

JCIM is now accepting images to appear on the front cover of the
journal. If you have a visually arresting and scientifically
interesting image, please email it and a brief (80 word) description
of the image to eic@jcim.acs.org. Additional information about the
technical specifications for the image can be found at
https://pubsapp.acs.org/paragonplus/submission/jcisd8/jcisd8_authguide.pdf.

Please contact the JCIM editorial office at eic@jcim.acs.org if you
have any specific questions regarding these guidelines.

JCIM has a Twitter feed (@JCIM_ACS; https://twitter.com/jcim_acs)
covering the latest news and highlights in the field. If you would
like us to feature your manuscript on this platform, please share a
Tweet of no more than 140 characters. Please feel free to share your
Twitter handle (if you have one) or any other relevant
department/institutional handles. Include this information in the
cover letter of your revised manuscript submission.

ACS Publications uses Crossref Similarity Check Powered by iThenticate
to detect instances of similarity in submitted manuscripts. In
publishing only original research, ACS is committed to deterring
plagiarism, including self-plagiarism. Your manuscript may be screened
for similarity to published material.

Sincerely,

Prof. Rommie Amaro
Associate Editor Journal of Chemical Information and Modeling
Urey Hall 3234, MS - 0340,
La Jolla, CA 92093-0340,
United States
Email : amaro-office@jcim.acs.org

------------------------------------ To revise your manuscript, log
into ACS Paragon Plus with your ACS ID at
http://paragonplus.acs.org/login and select "My Authoring Activity."
Your manuscript will be listed under "Revisions and Resubmissions
Requested by Editorial Office." Click on "Submit Revision" to begin
the submission process. Your manuscript number will be appended to
denote a revision.

When submitting your revised manuscript, you will be able to respond
to the comments made by the reviewer(s) in the text box provided or by
attaching a file containing your response letter.

------------------------------------

Reviewers' Comments to Author:

Reviewer: 1

Comments: Van der Spoel and co-workers built an atomistic model of the
cytoplasm of Escherichia coli composed of 1.5 million atoms and
investigated its behavior based on 3x1 us of molecular dynamics
(MD) simulations. Some interesting observations were made. The most
important aspects of this work are that i) the authors demonstrated
that it has become possible to perform MD simulations of cytoplasms
involving millions of atoms on the microsecond timescale, and ii) they
provide a Python script and a set of files that enable other
researchers to build their own E. coli cytoplasm models.

The work is of highest importance and novelty. Moreover, the
manuscript is nicely written and easy to follow. My recommendation
thus is to accept this manuscript for publication after following
minor comments have been addressed:

ANSWER: Thanks for the encouraging commments.

1) Why did the authors choose K+ and not Na+ to model the ionic
strength of 150 mmol/L? Please comment.

ANSWER: Like most organisms E. coli has a higher concentration of K+
than Na+ of 30-300 mmol/L versus 10 mmol/L Na+. The number 150 is
halfway the interval. A reference to a text book is added.

2) Figure 1A: Even though it is not of importance for the message
conveyed by this figure, this reviewer nonetheless wonders what the
difference between the red and green particles is in that figure. A
short note on it in the caption would be appreciated.

ANSWER: Red is water and green is biomolecules, this has been added to
the caption.

3) p. 13-15: It is mentioned that the RMSD of TufA, MetE and IcdA is
larger in the cytoplasm than in diluted solutions. Based on the RMSF
they also identified which part of these proteins become more
flexible. The authors later speculate that this is due to some
nonspecific interactions. Here, some more analysis is required. The
authors should demonstrate that these changes indeed arise from such
kind of interactions and not more specific interactions. And why are
only certain loops of each protein involved while all other loops seem
almost unaffected. And why does the RMSF of residues 450-460 of MetE
decrease?

ANSWER:

4) Figure 5: Why are the standard errors of the RMSF for CspC so much
larger than for the other systems. Please comment.

ANSWER:

5) In the discussion on p. 17 about the challenge of reproducing
protein-protein interactions one might also cite
DOI:10.1021/acs.jctc.8b00579 where it was shown that current force
fields have problems to distinguish between peptides known from
experiment to aggregate or not to aggregate. Moreover, increasing the
protein-water interactions as done in Amber03-ws may abolish protein
aggregation almost completely.

ANSWER: This paper was cited as well. Nevertheless, our earlier study
(Zhang et al, JCIM 58, 2018, 1037) showed that the DeltaGhydration are
slightly better using Amber99-sbws than Amber03-ws which is why we
used particular combination.

6) Following spelling/grammatical mistakes should be corrected:
p. 8, l. 29: A verb is missing in that sentence.
p. 10, l. 47: Change "don't" to "do not". =20
p. 11, l. 17: Add a space between "500" and "ps".
p. 15, l. 54: Change "of avoid" to "to avoid".
p. 16, l. 50: Change "duffusion" to "diffusion".

ANSWER: All fixed, thanks for finding these.

Reviewer: 2

Comments: Bortot et al. studied protein crowded systems using
molecular simulations with explicit solvent. MD simulations of
macromolecules in crowded conditions help us to understand protein
behaviors in a living cell. Recently, several computational scientists
performed MD simulations of multiple proteins in crowded conditions
and examined the differences of protein structure and dynamics in
crowded condition compared with that in dilute solution. In spite of
the importance of the research field, the paper has three critical
issues. At first, they used only small number of proteins and
tRNA(Phe) in the simulation system, although they call the set up as
the E. coli cytoplasm. Since the proteins can cover only 50 % of the
fraction of E. coli cytoplasm, it is difficult to call the system as
the E. coli cytoplasm. It should be called as a crowded condition that
mimic 50% of volume fraction of the cytosol of E. coli.

ANSWER:

They did not include ribosomes in the system. However, the size
differences between ribosomes and other proteins should be essential
in the cellular crowding effect.

ANSWER:

Secondly, they observed tRNA
aggregation with Mg2+, ATP, and FBP in their systems. They concluded
that completely protonating the phosphate groups of ATP and FBP is
necessary to prevent their aggregation with Mg2+. However, there is no
rationale or experimental evidences on the protonation state of
ATP. Presumably, the Mg2+ model or phosphate model may have
problems. Another possibility is that they used Amber99sb-ws, where
the 1.10 scaling is used for LJ interactions between water and
macromolecules. If they applied the scaling also to tRNA, tRNA
structure could be unfolded. However, there is no data about tRNA
structural changes in crowded condition compared to that in
solution.

ANSWER:

Thirdly, except for the protonation state, there are few novel
biophysical insights in the current work. Essentially same data on
diffusion or RMSD were already shown in the previous crowding papers,
although the length of simulations and the force fields are
different. Although the modeling strategy (Figure 1) seems to be
interesting, the rest of scientific results and discussion may have
rooms for improvements. In particular, the issue of the aggregation of
tRNA in the simulations should be overcome by the refinement of ion
models or the tRNA-water interaction scaling factors.

ANSWER: