### 2019-12-03 paper/rebuttal.txt
Updated rebuttal

Dear Prof. Amaro,

we were happy to receive constructive comments on our manuscript. We
have edited the manuscript according to the comments, and answered all
the questions. Please find a detailed list of replies below.

I hope the manuscript is acceptable in the present form and await your
decision.

Best regards,

David van der Spoel,
Uppsala University
Sweden.

MANUSCRIPT
- Please remove "et al." from your references and list all authors' names.
ANSWER: Done.

- Update the publication information in Reference 56 and Reference 73 when
available.
ANSWER: Done.

- Please include a Table of Contents graphic on the very last page of
your manuscript.  This graphic, in conjunction with the manuscript
title, should give the reader a summary of the research described in
the paper. The Table of Contents graphic will also be published in the
Abstract of the manuscript. This graphic can be no wider than 8 cm and
no taller than 4 cm and should be prepared following the
specifications under Artwork. Text should be limited to essential
technical labels.  More guidance is available at
http://pubs.acs.org/paragonplus/submission/toc_abstract_graphics_guidelines.pdf.
ANSWER: Done.

- The title page of your manuscript file needs more complete
 affiliation information for all authors. Could you please provide
 updated affiliations including all of the following information
 (where possible)? Department and institute (or company name), city,
 state/region, postal code, country.  Please indicate the
 corresponding author's email address with an asterisk.
ANSWER: It is complete.

- Funding Sources: Authors are required to report ALL funding sources
  and grant/award numbers relevant to this manuscript. Enter all
  sources of funding for ALL authors relevant to this manuscript in
  BOTH the Open Funder Registry tool in ACS Paragon Plus and in the
  manuscript to meet this requirement. See
  http://pubs.acs.org/page/4authors/funder_options.html for complete
  instructions.
ANSWER: Funding information was updated.

- ORCID: Authors submitting manuscript revisions are required to
  provide their own validated ORCID iDs before completing the
  submission, if an ORCID iD is not already associated with their ACS
  Paragon Plus user profiles. This iD may be provided during original
  manuscript submission or when submitting the manuscript
  revision. You can provide only your own ORCID iD, a unique
  researcher identifier. If your ORCID iD is not already validated and
  associated with your ACS Paragon Plus user profile, you may do so by
  following the ORCID-related links in the Email/Name section of your
  ACS Paragon Plus account. All authors are encouraged to register for
  and associate their own ORCID iDs with their ACS Paragon Plus
  profiles. The ORCID iD will be displayed in the published article
  for any author on a manuscript who has a validated ORCID iD
  associated with ACS Paragon Plus when the manuscript is
  accepted. Learn more at http://www.orcid.org.
ANSWER: We have added one ORCID.

- On resubmission you must submit the following:
a) A point-by-point response to the reviewers.
b) A clean, unmarked copy of the manuscript document file.
c) A marked copy of the revised manuscript in which all the changes are clearly tracked or highlighted. Upload this file as Supporting Information for Review.
d) A clean, unmarked Supporting Information for Publication file, without any highlighting or tracked changes.

- The reviewers indicate that the quality of the written English could
  be improved. Please consider using a native English speaker or an
  English-language editing service to assist with this. The ACS
  maintains a list of editing services:
  https://authoringservices.acs.org/.
ANSWER: We carefully went over the manuscript text and fixed errors.

Please contact the JCIM editorial office at eic@jcim.acs.org if you
have any specific questions regarding these guidelines.
ANSWER: We did not manage to make a cover image unfortunately.

Reviewers' Comments to Author:

Reviewer: 1

Comments: Van der Spoel and co-workers built an atomistic model of the
cytoplasm of Escherichia coli composed of 1.5 million atoms and
investigated its behavior based on 3x1 us of molecular dynamics
(MD) simulations. Some interesting observations were made. The most
important aspects of this work are that i) the authors demonstrated
that it has become possible to perform MD simulations of cytoplasms
involving millions of atoms on the microsecond timescale, and ii) they
provide a Python script and a set of files that enable other
researchers to build their own E. coli cytoplasm models.

The work is of highest importance and novelty. Moreover, the
manuscript is nicely written and easy to follow. My recommendation
thus is to accept this manuscript for publication after following
minor comments have been addressed:

ANSWER: Thanks for the encouraging commments.

1) Why did the authors choose K+ and not Na+ to model the ionic
strength of 150 mmol/L? Please comment.

ANSWER: Like most organisms E. coli has a higher concentration of K+
than Na+ of 30-300 mmol/L versus 10 mmol/L Na+. The number 150 is
halfway the interval. A reference to a text book is added.

2) Figure 1A: Even though it is not of importance for the message
conveyed by this figure, this reviewer nonetheless wonders what the
difference between the red and green particles is in that figure. A
short note on it in the caption would be appreciated.

ANSWER: Red is water and green is biomolecules, this has been added to
the caption.

3) p. 13-15: It is mentioned that the RMSD of TufA, MetE and IcdA is
larger in the cytoplasm than in diluted solutions. Based on the RMSF
they also identified which part of these proteins become more
flexible. The authors later speculate that this is due to some
nonspecific interactions. Here, some more analysis is required. The
authors should demonstrate that these changes indeed arise from such
kind of interactions and not more specific interactions. And why are
only certain loops of each protein involved while all other loops seem
almost unaffected. And why does the RMSF of residues 450-460 of MetE
decrease?

ANSWER: As explained in the text, the region comprised by residues
40-62 of TufA is a conformational switch that is known to be flexible,
residues 450-460 of MetE are a flexible loop and residues 1-8 of IcdA
form a flexible coil. Thus, one can expect these regions to display
variable conformational dynamics under different conditions. We
removed the speculation about the unspecific interactions and the
discussion about the RMSF difference (please see the answer to the
next question for details about this).

4) Figure 5: Why are the standard errors of the RMSF for CspC so much
larger than for the other systems. Please comment.

ANSWER: Thank you for pointing that out. There was a mistake in the
statistics of our RMSF calculations.  We corrected it and updated the
RMSF plots. The errors are much smaller now and the overall shift
towards higher RMSF values for the protein cytoplasm models is no
longer present. Accordingly, the following sentences were removed from
the text:

"In most cases (TufA, MetE, AhpC, CspC and Ppa), the residue-wise RMSF
values are higher in the cytoplasm model than under diluted
conditions, which indicates that proteins are more flexible in the
cytoplasm model (Fig.~\ref{fig:rmsf}). This is in agreement with
computational and experimental data showing that nonspecific
interactions with crowders can lead to destabilization of
proteins~\cite{Feig2011, miklos2011, Wang2012b}." from the results
section.

" and the flexibility profiles calculated considering the C$\upalpha$
for each chain of the elements of our cytoplasm model show that, in
the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC
and Ppa) are more flexible than in the dilute condition
(Fig.~\ref{fig:rmsf})." from the conclusion section.

"and \ref{fig:rmsf})" was removed from "the proteins are slightly less
stable in the cytoplasm model than in dilute condition (Figures
\ref{fig:structural_integrity_chain} and \ref{fig:rmsf})." in the
conclusion section

5) In the discussion on p. 17 about the challenge of reproducing
protein-protein interactions one might also cite
DOI:10.1021/acs.jctc.8b00579 where it was shown that current force
fields have problems to distinguish between peptides known from
experiment to aggregate or not to aggregate. Moreover, increasing the
protein-water interactions as done in Amber03-ws may abolish protein
aggregation almost completely.

ANSWER: This paper is now cited as well. Nevertheless, our earlier study
(Zhang et al, JCIM 58, 2018, 1037) showed that the DeltaGhydration are
slightly better using Amber99-sbws than Amber03-ws which is why we
used particular combination.

6) Following spelling/grammatical mistakes should be corrected:
p. 8, l. 29: A verb is missing in that sentence.
p. 10, l. 47: Change "don't" to "do not". =20
p. 11, l. 17: Add a space between "500" and "ps".
p. 15, l. 54: Change "of avoid" to "to avoid".
p. 16, l. 50: Change "duffusion" to "diffusion".

ANSWER: All fixed, thanks for finding these.

Reviewer: 2

Comments: Bortot et al. studied protein crowded systems using
molecular simulations with explicit solvent. MD simulations of
macromolecules in crowded conditions help us to understand protein
behaviors in a living cell. Recently, several computational scientists
performed MD simulations of multiple proteins in crowded conditions
and examined the differences of protein structure and dynamics in
crowded condition compared with that in dilute solution. In spite of
the importance of the research field, the paper has three critical
issues. At first, they used only small number of proteins and
tRNA(Phe) in the simulation system, although they call the set up as
the E. coli cytoplasm. Since the proteins can cover only 50 % of the
fraction of E. coli cytoplasm, it is difficult to call the system as
the E. coli cytoplasm. It should be called as a crowded condition that
mimic 50% of volume fraction of the cytosol of E. coli.

ANSWER: Obviously our model is not complete, to reflect this we
changed the title of the paper to "Making Soup: Preparing and
Validating Models of the Bacterial Cytoplasm for Molecular
Simulation". In the abstract the following sentence is added "The
model consists of 8 different proteins representing about 50% of the
cytoplasmic proteins and one type of t-RNA molecule."

They did not include ribosomes in the system. However, the size
differences between ribosomes and other proteins should be essential
in the cellular crowding effect.

ANSWER: This was a deliberate omission. A sentence was added to
explain it: "It has been estimated that a {\em E. coli} cell contains
about 50,000 ribosomes per $\mu^3$~\cite{Milo2017a}. The volume of our
box is about 12,000 nm$^3$ (\figref{shrink}), therefore in principle
one ribosome molecule should be added, however to reduce the
complexity this was omitted."

Secondly, they observed tRNA
aggregation with Mg2+, ATP, and FBP in their systems. They concluded
that completely protonating the phosphate groups of ATP and FBP is
necessary to prevent their aggregation with Mg2+. However, there is no
rationale or experimental evidences on the protonation state of
ATP. Presumably, the Mg2+ model or phosphate model may have
problems. Another possibility is that they used Amber99sb-ws, where
the 1.10 scaling is used for LJ interactions between water and
macromolecules. If they applied the scaling also to tRNA, tRNA
structure could be unfolded. However, there is no data about tRNA
structural changes in crowded condition compared to that in
solution.

ANSWER: There may be a number of possible solutions to the issue. The
work by Leontyev et al. (Refs 78 and 79) offers a plausible explanation
of the problem, namely that in fixed-charge models, that is
non-polarizable models, the infinite frequency dielectric constant
should be incorporated, either by introducing a relative dielectric
constant eps(inf) of approximately 1.7, or by scaling the charges down
by a factor of sqrt(1.7). Leontyev and co-workers advocate scaling the
charges since eps(inf) is somewhat material dependent. Such a solution
was described in the manuscript already, but a new sentence was added:
"Due to the omission of the electronic polarization, the screening is
underestimated and this corresponds to applying an effective
dielectric constant of about 1.7, alternatively scale the charges down
by $\sqrt{1.7}$ which would mean that FBP would get a charge of
approximately -3 instead of -4."

Thirdly, except for the protonation state, there are few novel
biophysical insights in the current work. Essentially same data on
diffusion or RMSD were already shown in the previous crowding papers,
although the length of simulations and the force fields are
different. Although the modeling strategy (Figure 1) seems to be
interesting, the rest of scientific results and discussion may have
rooms for improvements. In particular, the issue of the aggregation of
tRNA in the simulations should be overcome by the refinement of ion
models or the tRNA-water interaction scaling factors.

ANSWER: We agree that the amount of new biophysical insights is
limited. However, the purpose of this paper was rather to pave the way
for further in-depth studies by collecting information and scripts
allowing others to build on this work. For instance, as suggested by
this referee, one could add ribosomes in larger models for the E. coli
cytoplasm. We therefore think the present paper is particularly
suitable for the journal of chemical information and modeling.